---
title: "Module 9"
author: "Emily Yang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r SE}
n <- 1000
mu <- 3.5
sigma <- 4
v <- rnorm(n, mu, sigma)
s <- sample(v, size = 30, replace = FALSE)
m <- mean(s)
m
sd <- sd(s)
sem <- sd(s)/sqrt(length(s))
lower <- m - qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci <- c(lower, upper)
ci
```

## Central Limit Theorem
Distribution of averages of independent and identically distributed random variables becomes normal as the sample size increases. 
```{r CLT}
lambda <- 14
n <- 10
pop_se <- sqrt(lambda/n)  # the estimated SE
pop_se

x <- NULL
for (i in 1:1000) {
    x[i] <- mean(rpois(n = n, lambda = lambda)) # Assume μ and σ2 both = λ
}
hist(x, breaks = seq(from = lambda - 4 * sqrt(lambda)/sqrt(n), to = lambda +
    4 * sqrt(lambda)/sqrt(n), length.out = 20), probability = TRUE)

curve(dnorm(x, 0, 1), -4, 4, ylim = c(0, 0.8))
z <- (x - lambda)/pop_se
hist(z, breaks = seq(from = -4, to = 4, length.out = 20), probability = TRUE,
    add = TRUE)

qqnorm(x)
qqline(x)

n <- 100
x <- NULL
for (i in 1:1000) {
    x[i] <- sum(rpois(n = n, lambda = lambda))
}
hist(x, breaks = seq(min(x), max(x), length.out = 20), probability = TRUE)
```

CTL: Distribution will be approximately normal, will be centered at the population mean, and will have a standard deviation roughly equal to the standard error of the mean <br>

## Challenge

Expect normality if n×π and n×(1−π)
```{r CI}
n <- 1000
x <- 856
phat <- x/n # sucess/sample
phat

n * phat
n * (1 - phat) # both greater than 5
pop_se <- sqrt((phat) * (1 - phat)/n)
curve(dnorm(x, mean = phat, sd = pop_se), phat - 4 * pop_se, phat + 4 * pop_se)
upper <- phat + qnorm(0.975) * pop_se
lower <- phat - qnorm(0.975) * pop_se
ci <- c(lower, upper)
polygon(cbind(c(ci[1], seq(from = ci[1], to = ci[2], length.out = 1000), ci[2]),
    c(0, dnorm(seq(from = ci[1], to = ci[2], length.out = 1000), mean = phat,
        sd = pop_se), 0)), border = "gray", col = "green")
abline(v = ci)
abline(h = 0)

```
# Small Sample Confidence Interval
T distribution - common that we don't have enough data about population. Allows for more scatter around the mean. 
```{r Small Sample}
n <- 1e+05
mu <- 3.5
sigma <- 4
x <- rnorm(n, mu, sigma)
sample_size <- 30
s <- sample(x, size = sample_size, replace = FALSE)
m <- mean(s)
m
sd <- sd(s)
sd
sem <- sd(s)/sqrt(length(s))
sem
lower <- m - qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci_norm <- c(lower, upper)
ci_norm # don't use

lower <- m - qt(1 - 0.05/2, df = sample_size - 1) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qt(1 - 0.05/2, df = sample_size - 1) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci_t <- c(lower, upper)
ci_t # use
```